---
title: "Pokémon in NYC"
author: "Yingyu Cao"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    fig_height: 6
    fig_width: 8
---

```{r setup, include=FALSE}
 # keep this chunk in your .Rmd file
 knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

This document includes 2 part of work: 

* data cleaning 

* demo of visualizes the Pokémon showing up around NYC from 2:00 pm to 2:30 pm in 3 days (from July 26, 2016 to July 28, 2016). 

The interactive map enables viewers to zoom in and check the Pokémon ID, name, type, appear time, and location information.

*Codes for the demo can be find [here](https://github.com/Yingyu-Cao/Pokemon_NYC)*

## 1. Data
The data used in this demo comes from 2 Kaggle dataset:

* [pokemon-spawns](https://www.kaggle.com/kveykva/sf-bay-area-pokemon-go-spawns) is the spatial information mined from Pokémon Go from July 26, 2016 to July 28, 2016

* [pogo](https://www.kaggle.com/netzuel/pokmon-go-dataset-15-generations) contains the information of all Pokémon available in Pokémon GO.

```{r read data}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lubridate)
library(sp)
library(geojsonio)

dat <- read_csv('./data/pokemon-spawns.csv')
dat2 <- read_csv('./data/pogo.csv')
```

## 2. Data cleaning
The original data contains 3-day spatial information of Pokémon showing up all around the world. Only the NYC data from 2:00 pm to 2:30 pm in the 3 days is used in this demo. The author did the following data cleaning:

* deal with missing values, which are dropped because of its small proportion 

* parse and convert date to machine-readable format

* filter out useless data in two dataset

* calculate appearing time, and rarity of Pokémon

* combine two datasets based on the Pokémon ID
```{r data cleaning}
# preselect a small region around New York
# because the for loop for maping lat and lng to states takes long
# convert time to machine-readable format

# location of NY (roughly)
loc = c(40.7,-73.9)   # center
r = c(0.3,0.4)        # radius


dat <- dat %>% 
  mutate('appear_intvl' = (.$encounter_ms-.$disppear_ms)/60000,
         'lat' = round(.$lat,digits = 3),
         'lng' = round(.$lng,digits = 3)) %>% 
  mutate('appear_time' = 
           format(as.POSIXct(.$encounter_ms/1000, 
                           origin="1970-01-01", 
                           tz="America/New_York"),"%H:%M")) %>% 
  filter(appear_intvl>0,
         appear_time <= format(strptime('14:30',"%H:%M"),"%H:%M"),
         lat>loc[1]-r[1],lat<loc[1]+r[1],
         lng>loc[2]-r[2],lng<loc[2]+r[2]) %>% 
  select(c('num','name','lat','lng','appear_intvl','appear_time'))
```

```{r map state info}
# read in usa state map information
usa <- geojson_read(
  "http://eric.clst.org/assets/wiki/uploads/Stuff/gz_2010_us_040_00_500k.json",
  what = "sp")

# map state info to each point by its lat and lng
for (i in 1:nrow(dat)) {
  coords <- c(dat$lng[i], dat$lat[i])
  if(any(is.na(coords))) next
  point <- sp::SpatialPoints(
    matrix(
      coords,
      nrow = 1
    )
  )
  sp::proj4string(point) <- sp::proj4string(usa)
  polygon_check <- sp::over(point, usa)
  dat$state[i] <- as.character(polygon_check$NAME)
}

# select points only in NY
dat <- dat %>% filter(state == 'New York')
```

```{r pokemon type}
# get pokemon type info from pogo dataset
dat2 <- dat2 %>% 
  mutate('num' = Pokedex) %>% 
  select(num,Primary)

# compute rarity and average appearing time
# combine location dataset with pogo dataset
# rarity calculation: sum up all the appearing time all over NY. 
# <30mins: rare, 30-60mins: uncommon, 60mins: common
dat <- dat %>% group_by(num,name,lat,lng) %>% 
  summarise('ave_appear_time' = sum(appear_intvl)/n(),
            'total_appear_time'= sum(appear_intvl)) %>% 
  mutate('freq'= 'uncommon',
         'freq' = ifelse(total_appear_time<30,'rare',freq),
         'freq' = ifelse(total_appear_time>60,'common',freq)) %>% 
  # filter(ave_appear_time>5) %>% 
  left_join(dat2,by='num')
```

```{r}
# no need for this chunk because all NY info all collected during 14:00-15:00
# # create breaks
# breaks <- hour(hm("12:00", "13:00", "14:00", "15:00", "16:00"))
# # labels for the breaks
# labels <- c("12:00-13:00", "13:00-14:00", "14:00-15:00", "15:00-16:00")
#   
# dat <- dat %>% mutate('appear_time2' = cut(x=hour(.$appear_time), 
#                                            breaks = breaks, 
#                                            labels = labels, 
#                                            include.lowest=TRUE)) %>% 
#   group_by(num,name,appear_time2,lat,lng) %>% 
#   summarise('ave_appear_time' = sum(appear_intvl)/n()) %>% 
#   filter(ave_appear_time>3)

```

## 3. Visualization
The map contains 1311 points, showing where each Pokémon appeared from 2:00 pm to 2:30 pm, from July 26, 2016 to July 28, 2016. Each wild Pokémon encountered is a particular point, with the size and color of the point represent how long a Pokémon appeared on average and its type. 
```{r}
# abadonded, find a better method
# library(maptools)
# library(maps)

# mapWorld <- borders("state", colour="gray50", fill="white")
# register_google(key = "AIzaSyBPwH5XIOe2yvySzbzLWUuLYDDfVr-7UNo")
# mymap <- get_map(location = c(lon = -74.00, lat = 40.77), zoom = 12, maptype = "roadmap")
# mymap <- ggmap(mymap)+
#   geom_point(data = dat, 
#              aes(x = lng, y = lat, 
#                  alpha = ave_appear_time, 
#                  col = appear_time2))
```

```{r visualization}
library(plotly)
library(RColorBrewer)
# need to be hidden
# here for more info https://docs.mapbox.com/help/how-mapbox-works/access-tokens/
# Sys.setenv("MAPBOX_TOKEN" = 
#              'pk.eyJ1IjoieXVyaWlpaSIsImEiOiJja2QyN3MxOW0wZzFyMndueWR4b2l3cnpnIn0.9BcXtfA3YWLSTXGV-3QxHw') # for Orca

mycolors = c(brewer.pal(name="Set1", n = 9), brewer.pal(name="Accent", n = 5))

# sizeref: smaller, larger points
fig <- dat %>% plot_ly(lat = ~lat,
                       lon = ~lng,
                       marker = list(size = ~ave_appear_time,
                                     sizeref = 0.2,
                                     sizemode = 'area'),
                       color = ~Primary,
                       colors = mycolors,
                       type = 'scattermapbox',
                       mode = "markers",
                       hoverinfo = 'text',
                       hovertext = ~paste('</br> ID:',num,
                                         '</br> Name:',name,
                                         '</br> Type:',Primary,
                                         '</br> Appear time:',
                                         round(ave_appear_time,digits = 0),'mins',
                                         '</br> Location:',lng,'°,',lat,'°'))
# center: make sure the map shows all the points when being loaded
# zoom: larger, more details
fig <- fig %>%
  layout(
    mapbox = list(
      # style = 'open-street-map',
      style = 'carto-positron',
      zoom =11,
      center = list(lon = loc[2]-r[2]/4, lat = loc[1]+r[1]/6)),
    title = 'Pokemon Distribution')

fig
```

```{r include=FALSE}
# this is the refined version of sp package and should be much faster, but something is wrong with the working space storage
# library(sf)
# library(spData)
# 
# ## pointsDF: A data.frame whose first column contains longitudes and
# ##           whose second column contains latitudes.
# ##
# ## states:   An sf MULTIPOLYGON object with 50 states plus DC.
# ##
# ## name_col: Name of a column in `states` that supplies the states'
# ##           names.
# lonlat_to_state <- function(pointsDF,
#                             states = spData::us_states,
#                             name_col = "NAME") {
#     ## Convert points data.frame to an sf POINTS object
#     pts <- st_as_sf(pointsDF, coords = 1:2, crs = 4326)
# 
#     ## Transform spatial data to some planar coordinate system
#     ## (e.g. Web Mercator) as required for geometric operations
#     states <- st_transform(states, crs = 3857)
#     pts <- st_transform(pts, crs = 3857)
# 
#     ## Find names of state (if any) intersected by each point
#     state_names <- states[[name_col]]
#     ii <- as.integer(st_intersects(pts, states))
#     state_names[ii]
# }
# 
# testPoints <- data.frame(x = c(-90, -120, 0), y = c(44, 44, 44))
# lonlat_to_state(testPoints)
# 
# write_csv(dat,'test.csv')
# dat <- read_csv('test.csv')
# 
# test <- dat %>% group_by(num,name,lat,lng) %>% 
#   summarise('ave_appear_time' = sum(appear_intvl)/n(),
#             'total_appear_time'= sum(appear_intvl))
# test <-  test[1:round(nrow(test)),c('lng','lat')]
# pts <- st_as_sf(test, coords = 1:2, crs = 4326)
# states = spData::us_states
# states <- st_transform(states, crs = 3857)

# lonlat_to_state(data.frame(x = dat$lng,y = dat$lat))
```


